---
title: "RiboSTAMP SRRM3 GSE314176 — results summary"
author: "SRF lab"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

suppressPackageStartupMessages({
  library(data.table)
  library(ggplot2)
  library(knitr)
})

BASE_DIR <- normalizePath(file.path(dirname(knitr::current_input()), ".."), mustWork = FALSE)
TABLES   <- file.path(BASE_DIR, "results", "tables")
FIGURES  <- file.path(BASE_DIR, "results", "figures")
```

## Deliverable 1 — Novel Srrm3 isoform in progenitor-like cells

**Short-read anchor** (from `90-1239779069/SRRM3_novel_exon/`):
`chr5:135,898,574–135,898,652` (mm39, 79 bp, negative strand). Bulk PSI: Parental 57% > Pos 27% > KO 5%.

### Per-cluster inclusion

```{r per-cluster}
pc_path <- file.path(TABLES, "novel_srrm3_per_cluster.tsv")
if (file.exists(pc_path) && file.info(pc_path)$size > 0) {
  per_cluster <- fread(pc_path)
  knitr::kable(per_cluster, digits = 3,
               caption = "Novel Srrm3 isoform detection per hippocampal cluster (P25).")
} else {
  cat("No per-cluster table yet — run 04_progenitor_enrichment.R.")
}
```

### Fisher test: progenitor vs other

```{r fisher}
fi_path <- file.path(TABLES, "novel_srrm3_fisher_enrichment.tsv")
if (file.exists(fi_path) && file.info(fi_path)$size > 0) {
  fi <- fread(fi_path)
  knitr::kable(fi, digits = 4)
} else {
  cat("No Fisher result yet.")
}
```

### Cluster PSI plot

```{r psi-plot, fig.width=9, fig.height=5}
psi_pdf <- file.path(FIGURES, "novel_srrm3_cluster_psi.pdf")
if (file.exists(psi_pdf)) {
  cat(sprintf("See `%s`.", psi_pdf))
}
```

## Deliverable 2 — Interneuron isoform list

```{r interneuron}
in_path <- file.path(TABLES, "interneuron_isoforms.tsv")
if (file.exists(in_path)) {
  ino <- fread(in_path)
  knitr::kable(head(ino, 30), digits = 3,
               caption = "Top 30 GABA-enriched isoforms (full table in results/tables/).")
  cat(sprintf("\nTotal isoforms: %d\n", nrow(ino)))
} else {
  cat("Run 05_interneuron_isoform_table.R + 06_editsc_translation_column.R.")
}
```

## Methods quick reference

- **Alignment:** pbmm2 `--preset ISOSEQ` to mm10 + Gencode vM25.
- **Discovery:** IsoQuant targeted to Srrm3 locus ± 10 kb (primary); full-tx on demand.
- **Classification:** SQANTI3.
- **Cross-build validation:** UCSC liftOver mm10 → mm39; bedtools intersect vs the 79-bp anchor.
- **Enrichment:** Fisher exact on 2×2 (detects novel × progenitor).
- **Interneurons:** author's annotated isoform matrix subsetted to GABA cells; translation via EditsC matrix.
